---
title: "KaylaWilliams_A4"
author: "Kayla Williams"
date: "10/12/2021"
output: html_document
---

```{r}
library(tidyverse)
library(censusapi)
Sys.setenv(CENSUS_KEY = "e7f4fdf1cb4024223aeaafafa8d8943523ddcefb")
```
```{r}
acs_vars_2019_5yr <-
  listCensusMetadata(
    name = "2019/acs/acs5",
    type = "variables"
  )
```

```{r}
race_categories = c("White Alone", "Black or African American Alone", "American Indian and Alaska Native Alone", "Asian Alone", "Native Hawaiian and Other Pacific Islander Alone", "Some Other Race Alone", "Some Other Race Alone")
bay_educational_attainment_race_2019 = 
  1:7 %>%
  #assign 1-7 to x
  map_dfr(function(x){
    
    getCensus(
  name = "acs/acs5", 
  vintage= "2019", 
  region = "county: 001, 013, 041, 055, 075, 081, 085, 095, 097",
  regionin = "state:06", 
  vars = paste0("group(C15002", LETTERS[x], ")")
) %>% 
  select(ends_with("E") &-c(state, NAME)) %>% 
  summarize_all(sum) %>% 
  pivot_longer(
    everything(),
    names_to = "name", 
    values_to ="estimate"
  ) %>% left_join(acs_vars_2019_5yr %>%
                    select(name,label)
                  )%>%
  separate(
    label,
    into = c(NA, NA, "sex", "education"),
    sep = "!!"
  ) %>%
  select(-name) %>%
  filter(!is.na(education)) %>%
      group_by(education) %>%
  summarise(estimate = sum(estimate)) %>%
  mutate(race= race_categories[x])
    
  })
```
```{r}
schoolorder = c("Bachelor's degree or higher", "Some college or associate's degree", "High school graduate (includes equivalency)", "Less than high school diploma")

bay_race_total = 
  bay_educational_attainment_race_2019 %>% 
  group_by(race) %>%
  summarise(estimate = sum(estimate)) %>%
  mutate(education = "Total")
```
```{r}
#PROPORTION 
bay_educational_attainment_race_2019 %>%
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>%
  rbind(bay_race_total) %>%
  ggplot() +
  geom_bar(
    aes(
      x= education %>% factor(levels = rev(c("Total", unique(schoolorder)))),
      y = estimate, 
      fill = race %>% factor(levels = rev(unique(bay_educational_attainment_race_2019$race)))
    ),  
      stat = "identity",
      position = "fill"
    ) + 
      labs(
        x = "Educational Attainment",
        y = "Number of Individuals",
        title = "Bay Area Educational Attainment by Race",
        fill = "Race"
      ) + 
      coord_flip() + 
  theme(legend.position = "bottom", 
        legend.direction = "vertical")
```
```{r}
#COUNTS
bay_educational_attainment_race_2019 %>%
  group_by(education, race) %>% 
  summarize(estimate = sum(estimate)) %>%
  ggplot() +
  geom_bar(
    aes(
      x= education %>% factor(levels = rev(unique(schoolorder))),
      y = estimate, 
      fill = race %>% factor(levels = rev(unique(bay_educational_attainment_race_2019$race)))
    ),  
      stat = "identity",
      position = "stack"
    ) + 
      labs(
        x = "Educational Attainment",
        y = "Number of Individuals",
        title = "Bay Area Educational Attainment by Race",
        fill = "Race"
      ) + 
      coord_flip() + 
  theme(legend.position = "bottom",
        legend.direction = "vertical")
  
```
 
 The two graphs, especially the "fill" version which highlights the
 proportions of races within each educational attainment level, as opposed to 
 raw counts, suggest that there is indeed inequity within education outcomes.
 
 In an ideally equal world, the breakdown of educational achievement by race 
 would mirror the proportions seen in the population breakdown by race. That is 
 to say that the percentages of each race in a bucket of schooling levels would 
 be nearly equal to that race's percentage of the population, as this would
 signify eqqual representation. However, the graph of proportions highlights that 
 Black people, for instance, are heavily overrepresented in the population of
 folks with less than a high school diploma and underrepresented in the 
 population of those with bachelor's degrees and higher as their fractions in
 these two buckets vary from their fraction of the overall population.
 
 White and Asian populations, in contrast, are slightly overrepresented in 
 those with higher degrees and white people are also pretty significantly 
 underrepresented amongst those without high school diplomas. 



